---
title: "Assignment 07"
output: html_document
date: '2022-07-26'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(readr)
library(dplyr)
library(car)
library(gvlma)
```


# Importing Data 

```{r}
library(dplyr)
Property <- read.csv("data/prop_prices_reduced.csv")

#Renaming column sale_def with an intuitive name: price

Property <- rename(Property, price = sale_def)

# Factor in  data: Pool
Property$pool <- factor(Property$pool)

```

# Separate data into testing and training

```{r}
set.seed(1234)
data = sort(sample(nrow(Property), nrow(Property)*.9))

#creating training data set by selecting the output row values
train<-Property[data,]
#creating test data set by not selecting the output row values
test<-Property[-data,]
```

# Run previous final model to the training data

```{r}
library(car)
train$logprice =log(train$price)
train$logarea= log(train$area)
train$logbath= log(train$bath)
train$logarea_heated = log(train$area_heated)
train$logDist_lakes = log(train$dist_lakes)
 
library(gvlma)
Fit_train <- lm(logprice ~ bed + logbath + area + logarea_heated + dist_cbd + logDist_lakes + pool, data = train)

gvmodeltrain<- gvlma(Fit_train)
summary(gvmodeltrain)

par(mfrow = c(2, 2))
plot(Fit_train)
```

# Run the standard LASSO regression model on the training data.
```{r}
# Factor pool
train$pool <- factor(train$pool)

# Generating dependent variable
y <- train$logprice

# Generate Independent Variables
library(Matrix)
x <- sparseMatrix (a =as.numeric(train$bed), b =as.numeric(train$logbath), c =as.numeric(train$logbath))

, d =as.numeric(train$area), e =as.numeric(train$logarea_heated), f =as.numeric(train$dist_cbd), g =as.numeric(train$logDist_lakes), H =as.numeric(train$pool))


library(gamlr)
LASSOmodel <- gamlr(x, y, verb=TRUE)

summary(LASSOmodel)
plot(LASSOmodel)
```

# Run a 10-fold cross-validated LASSO on the training data
```{r}
cv.LASSOmodel <- cv.gamlr(x, y, verb=TRUE,nfold =10)

```


# Calculate the RMSE for each of the lambda's selection methods discussed (AIC, BIC, AICc, cv.min, cv.1se) using the testing data. Which method performed the best in prediction the home price?
```{r}
log(LASSOmodel$lambda[which.min(AICc(LASSOmodel))])
log(LASSOmodel$lambda[which.min(AIC(LASSOmodel))])
log(LASSOmodel$lambda[which.min(BIC(LASSOmodel))])
log(cv.spender$lambda.min)
log(cv.spender$lambda.1se)

```
